---
title: "Feature Engineering"
output: html_document
---
#### Data Cleaning & Processing
This section checks missing values, creates 6 new features (e.g. tour conversion rate, tour booked every sellable capacity, etc), and 
checks data abnormality. 
```{r}
#check for missing values (no missing values)
tidyr::gather(summarize_each(data_labeled, funs(sum(is.na(.))/n())), key = "feature", value = "missing_pct")
data_labeled <- na.omit(data_labeled)  #drop NAs if any

#create new features
data_labeled <- dplyr::mutate(data_labeled, 
                      tour_conversion = tour_completed / ifelse(tour_booked == 0, 1, tour_booked),
                      tour_booked_per_sellable_capacity = tour_booked / ifelse(sellable_capacity == 0, 1, sellable_capacity),
                      tour_completed_per_sellable_capacity = tour_completed / ifelse(sellable_capacity == 0, 1, sellable_capacity),
                      sales_time_per_desk = total_sales_time / ifelse(desks_sold ==0, 1, desks_sold),
                      avg_signup_per_event = total_attendee / ifelse(total_event == 0, 1, total_event),
                      event_participation_rate = total_attendee / ifelse(occupied_capacity == 0, 1, occupied_capacity))
head(data_labeled)

## check data abnormality
summary(data_labeled)   
##     issue 1: event participation shouldn't be a negative number or anything too high (>100)
data_labeled <- data_labeled[data_labeled$event_participation_rate >= 0 | data_labeled$event_participation_rate <= 100,]  
##     issue 2: tour conversion rate is a percentage so it shouldn't go over 1
data_labeled <- data_labeled[data_labeled$tour_conversion < 2, ]
##     issue 3: shouldn't be any is_mature 
data_labeled <- data_labeled[data_labeled$is_mature != -1, ]
##     issue 4: median lead time abnormality
data_labeled <- data_labeled[!(data_labeled$median_lead_time > 100 & data_labeled$occ < 0.8), ]
##     issue 5: ac_issue_count abnormality (data accuracy)
data_labeled <- data_labeled[data_labeled$ac_issue_count <= 1000, ]
dim(data_labeled)  #9431 x 49
```

#### Numerical Variables Check
Before we build the model, we want to take a look at the distribution of all numeric variables for the purposes of initial selection.
```{r}
## check numerical variable distribution
#### category 1: year, month, performance variables
data_labeled %>%
  select(year, month, net_arpm, gross_arpm, occ, financial_occ, discount) %>% 
  tidyr::gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_density()

##### category 2: market demand variables
data_labeled %>% 
  select(median_lead_time, rush_booking_pct, tour_booked_per_sellable_capacity, tour_booked_per_sellable_capacity, tour_conversion,
         sales_time_per_desk, churn_rate) %>% 
  tidyr::gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_density()

##### category 3: Zendesk variables
data_labeled %>% 
  select(printing_issue_count, ac_issue_count, billing_issue_count, maintenance_issue_count, printing_issue_solve_speed, ac_issue_solve_speed, billing_issue_solve_speed, maintenance_issue_solve_speed) %>% 
  tidyr::gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_density()

#### category 4: event variables
#### community event, office hour event, orientation event, other_event might not be good predictors
data_labeled %>% 
  select(happy_hour_event, tgim_event, business_development_event, food_for_thought_event, office_hour_event, community_event, other_event, total_event, total_attendee, orientation_event, wellness_event, total_event, business_development_event, avg_signup_per_event, event_participation_rate) %>% 
  tidyr::gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_density()
```


#### Feature Engineering
This section details the steps of feature engineering: turning the character-type metrics into factors, taking 3-month's moving average of metrics, and sanity-checking data. 

The idea of taking moving average of all the metrics is to create a window for prediction. This makes sense from both a business and a technical perspective. From the business perspective, the hypothesis is that the future occupancy is highly dependent on how the building and market performed the past 3 months. From the technical perspective, a 3-month window is needed so that the prediction could be performed ahead of time.
```{r}
#-------------feature engineering--------------
#### 1. change character into factor type
data_labeled$name <- as.factor(data_labeled$name)
data_labeled$year <- as.factor(data_labeled$year)
data_labeled$month <- as.factor(data_labeled$month)
data_labeled$city <- as.factor(data_labeled$city)
data_labeled$market <- as.factor(data_labeled$market)
data_labeled$region <- as.factor(data_labeled$region)
data_labeled$is_mature <- as.factor(data_labeled$is_mature)
head(data_labeled)

#### 2. moving average
model_data <- data_labeled %>% 
  dplyr::group_by(name) %>% 
  dplyr::mutate(historical_occupancy = rollapply(occ, 3, mean, align = 'right', fill = NA),
                performance_cluster = rollapply(performance_cluster, 3, mean, align = 'right', fill = NA),
                median_lead_time = rollapply(median_lead_time, 3, mean, align = 'right', fill = NA),
                rush_booking_pct = rollapply(rush_booking_pct, 3, mean, align = 'right', fill = NA),
                tour_conversion = rollapply(tour_conversion, 3, mean, align = 'right', fill = NA),
                tour_booked = rollapply(tour_booked, 3, mean, align = 'right', fill = NA),
                tour_completed = rollapply(tour_completed, 3, mean, align = 'right', fill = NA),
                desks_sold = rollapply(desks_sold, 3, mean, align = 'right', fill = NA),
                total_sales_time = rollapply(total_sales_time, 3, mean, align = 'right', fill = NA),
                tour_booked_per_sellable_capacity = rollapply(tour_booked_per_sellable_capacity, 3, mean, align = 'right', fill = NA),
                tour_completed_per_sellable_capacity = rollapply(tour_completed_per_sellable_capacity, 3, mean, align = 'right', fill = NA),
                sales_time_per_desk = rollapply(sales_time_per_desk, 3, mean, align = 'right', fill = NA),
                churn_rate = rollapply(churn_rate, 3, mean, align = 'right', fill = NA),
                printing_issue_count = rollapply(printing_issue_count, 3, mean, align = 'right', fill = NA),
                ac_issue_count = rollapply(ac_issue_count, 3, mean, align = 'right', fill = NA),
                billing_issue_count = rollapply(billing_issue_count, 3, mean, align = 'right', fill = NA),
                maintenance_issue_count = rollapply(maintenance_issue_count, 3, mean, align = 'right', fill = NA),
                printing_issue_solve_speed = rollapply(printing_issue_solve_speed, 3, mean, align = 'right', fill =NA),
                ac_issue_solve_speed = rollapply(ac_issue_solve_speed, 3, mean, align = 'right', fill = NA),
                billing_issue_solve_speed = rollapply(billing_issue_solve_speed, 3, mean, align = 'right', fill = NA),
                maintenance_issue_solve_speed = rollapply(maintenance_issue_solve_speed, 3, mean, align = 'right', fill = NA),
                other_event = rollapply(other_event, 3, mean, align = 'right', fill = NA),
                business_development_event = rollapply(business_development_event, 3, mean, align = 'right', fill = NA),
                happy_hour_event = rollapply(happy_hour_event, 3, mean, align = 'right', fill = NA),
                tgim_event = rollapply(tgim_event, 3, mean, align = 'right', fill = NA),
                food_for_thought_event = rollapply(food_for_thought_event, 3, mean, align = 'right', fill = NA),
                office_hour_event = rollapply(office_hour_event, 3, mean, align = 'right', fill = NA),
                orientation_event = rollapply(orientation_event, 3, mean, align = 'right', fill = NA),
                wellness_event = rollapply(wellness_event, 3, mean, align = 'right', fill = NA),
                avg_signup_per_event = rollapply(avg_signup_per_event, 3, mean, align = 'right', fill = NA),
                total_event = rollapply(total_event, 3, mean, align = 'right', fill = NA),
                total_attendee = rollapply(total_attendee, 3, mean, align = 'right', fill = NA),
                event_participation_rate = rollapply(event_participation_rate, 3, mean, align = 'right', fill = NA)) %>% 
  select(-c("net_arpm", "gross_arpm", "financial_occ", "sellable_capacity", "total_capacity", "occupied_capacity", "community_event", "orientation_event", "office_hour_event", "other_event", "tour_booked", "tour_completed", "desks_sold", "total_sales_time", "total_event", "total_attendee"))

model_data <- na.omit(model_data)  #drop NA's from taking moving average
dim(model_data)                    #8435, 34

#### 3. sanity-check model_data
head(model_data)
table(model_data$performance_cluster)  #check averaged performance cluster
#write.csv(model_data, "model_final_data.csv")   #only write data when necessary
```

##### correlation analysis
Before we move onto the modeling part, we want to get a quick look at the correlations of occupancy and features as an initial selection. I divided up the variables into sales, tour, event, and zendesk categories and used the correlation matrices to visualize variable correlation.

From the correlation analysis, median_lead_time, rush_booking_pct, desks_sold, tour_booked_per_sellable_capacity, tour_conversion, event_participation_rate, avg_sign_up_per_event, maintenance_issue_count, and billing_issue_solve_speed are correlated to occupancy. The matrices also identified highly-correlated variable pairs, including tour_booked_per_sellable_capacity and tour_completed_per_sellable_capacity.
```{r}
#### important sales variables
new_data <- model_data %>% filter(date > '2018-01-01')
sales_matrix <- as.matrix(model_data[c("occ","discount", "median_lead_time", "rush_booking_pct", "sales_time_per_desk", "churn_rate")])

sales_matrix %>% 
  cor() %>% 
  round(3) %>% 
  ggcorrplot(lab = T, 
             colors = c("#6D9EC1", "white", "#E46726"),
             outline.col = "white",
             ggtheme = ggplot2::theme_gray) + 
             ggtitle('Occupancy vs. Sales Metrics')

#### important tour variables
tour_matrix <- as.matrix(model_data[c("occ", "tour_booked_per_sellable_capacity", "tour_completed_per_sellable_capacity", "tour_conversion")])

tour_matrix %>% 
  cor() %>% 
  round(3) %>% 
  ggcorrplot(lab = T, 
             colors = c("#6D9EC1", "white", "#E46726"),
             outline.col = "white",
             ggtheme = ggplot2::theme_gray) + 
             ggtitle('Occupancy vs. Tour Metrics')

#### important event variables
event_matrix <- as.matrix(new_data[c("occ","tgim_event",  "wellness_event", "avg_signup_per_event", "event_participation_rate")])

event_matrix %>% 
  cor() %>% 
  round(3) %>% 
  ggcorrplot(lab = T, 
             colors = c("#6D9EC1", "white", "#E46726"),
             outline.col = "white",
             ggtheme = ggplot2::theme_gray) + 
             ggtitle('Occupancy vs. Event Metrics')

#### important zendesk variable
zendesk_matrix <- as.matrix(new_data[c("occ", "ac_issue_count","maintenance_issue_count", "printing_issue_solve_speed","billing_issue_solve_speed", "printing_issue_count" )])

zendesk_matrix %>% 
  cor() %>% 
  round(3) %>% 
  ggcorrplot(lab = T, 
             colors = c("#6D9EC1", "white", "#E46726"),
             outline.col = "white",
             ggtheme = ggplot2::theme_gray) + 
             ggtitle('Occupancy vs. Zendesk Metrics')
```

